\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-05-19 paper.tex
update stopping rule

\begin{algorithm}[!htb]
\DontPrintSemicolon
\caption{QNMCEM algorithm for the lights model inference}
\label{alg:QNMCEM}
\KwInput{Training data $\cD_n$; initialize $N=50L$; tuning hyper-parameters $(\zeta_{1,k}, \zeta_{2,k}, \zeta_{3,k})_{k=0,\ldots,K-1}$}
\KwOutput{Last parameters $\hat \theta \in \R^\vartheta$}
\tcc{Initialization}
Compute the starting parameters $\theta^{(0)} \in \R^\vartheta$ \\
\For{$w = 1,\ldots,$ until convergence}    
{ 
  \tcc{Monte Carlo E-step}
  Compute $\hat \Lambda_{ik,\tilde g_i}^{\theta^{(w)}}$ using~\eqref{eq:hat_Lambda_ik} for the required functions $\tilde g_i$\\
  Compute $\hat \pi_{ik}^{\theta^{(w)}}$ using~\eqref{eq:hat_pi_ik-def}\\
  \If{Condition}
  {
      $N \leftarrow N + $\\
      $tol \leftarrow tol - $
  }
  \tcc{Quasi-Newton M-step}
  Update $D^{(w+1)}$ using~\eqref{eq:update_D}\\
  Update $(\xi_k^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:sub-problem-xi_k} \\
  Update $(\beta_k^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:beta-update} \\
  Update $({\gamma_k^0}^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:sub-problem-gamma_k^0} \\
  Update $(\gamma_k^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:gamma-update} \\
  Update $\lambda_0^{(w+1)}$ using~\eqref{eq:update_lambda_0}\\
  Update $\phi^{(w+1)}$ using~\eqref{eq:update_phi} \\
}
\textbf{Return:} {$\hat \theta$}
\end{algorithm}
The convergence condition in Algorithm~\ref{alg:QNMCEM} is cv$(\Delta_{\text{rel}}^{(w+1)}) >$ cv$(\Delta_{\text{rel}}^{(w)})$, where cv$(\Delta_{\text{rel}}^{(w+1)})$ represents the variation at the $(w + 1)$-th iteration computed as
\begin{equation*}
  \text{cv}(\Delta_{\text{rel}}^{(w+1)})= \dfrac{\text{sd}(\Delta_{\text{rel}}^{(w-1)}, \Delta_{\text{rel}}^{(w)}, \Delta_{\text{rel}}^{(w+1)})}{\text{mean}(\Delta_{\text{rel}}^{(w-1)}, \Delta_{\text{rel}}^{(w)}, \Delta_{\text{rel}}^{(w+1)})},
\end{equation*}
where sd(.) and mean(.) are the empirical standard deviation and mean functions respectively, and relative difference $\Delta_{\text{rel}}^{(w+1)}$ is given by 
\begin{equation*}
  \label{eq:relative_difference}
  \Delta_{\text{rel}}^{(w+1)} = \text{max} \Big\{\dfrac{|\theta^{(w+1)} - \theta^{(w)}|}{|\theta^{(w)}| + \epsilon } \Big\}.
\end{equation*}

>>>>

\begin{algorithm}[!htb]
\DontPrintSemicolon
\caption{QNMCEM algorithm for the lights model inference}
\label{alg:QNMCEM}
\KwInput{Training data $\cD_n$; initialize $N=50L$; tuning hyper-parameters $(\zeta_{1,k}, \zeta_{2,k}, \zeta_{3,k})_{k=0,\ldots,K-1}$}
\KwOutput{Last parameters $\hat \theta \in \R^\vartheta$}
\tcc{Initialization}
Compute the starting parameters $\theta^{(0)} \in \R^\vartheta$ \\
\For{$w = 1,\ldots,$ until convergence}    
{ 
  \tcc{Monte Carlo E-step}
  Compute $\hat \Lambda_{ik,\tilde g_i}^{\theta^{(w)}}$ using~\eqref{eq:hat_Lambda_ik} for the required functions $\tilde g_i$\\
  Compute $\hat \pi_{ik}^{\theta^{(w)}}$ using~\eqref{eq:hat_pi_ik-def}\\
  \If{Condition}
  {
      $N \leftarrow N + $\\
      $tol \leftarrow tol - $
  }
  \tcc{Quasi-Newton M-step}
  Update $D^{(w+1)}$ using~\eqref{eq:update_D}\\
  Update $(\xi_k^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:sub-problem-xi_k} \\
  Update $(\beta_k^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:beta-update} \\
  Update $({\gamma_k^0}^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:sub-problem-gamma_k^0} \\
  Update $(\gamma_k^{(w+1)})_{k=0,\ldots,K-1}$ by solving~\eqref{eq:gamma-update} \\
  Update $\lambda_0^{(w+1)}$ using~\eqref{eq:update_lambda_0}\\
  Update $\phi^{(w+1)}$ using~\eqref{eq:update_phi} \\
}
\textbf{Return:} {$\hat \theta$}
\end{algorithm}
Let us define the relative difference at step $(w+1)$ as
\[\Delta_{\text{rel}}^{(w+1)} = \max_{j=1,\ldots,\vartheta} \Big\{\dfrac{|\theta^{(w+1)}_j - \theta^{(w)}_j|}{|\theta^{(w)}_j| + \epsilon_1 } \Big\}.\]
Then in practice, the convergence for Algorithm~\ref{alg:QNMCEM} is declared when the condition
\[\Delta_{\text{rel}}^{(w+1)} < \epsilon_2 \]
is satisfied for 3 consecutive iterations, since the monotonicity property is lost due to the Monte Carlo approximations.
